---
title: "assignment2"
output: pdf_document
date: "2023-02-15"
---

Running the file provided by Professor first and then starting with assignment 02

```{r}
#HW02.R
#run a lasso and elastic net model
library(survival)
library(coxed)
library(glmnet)
library(corrplot)
```


```{r}
#Simulate time-to-event data
set.seed(123)
n <- 1000

x1 = matrix(rbinom(n * 10, size = 1, prob = 0.3), n, 10)

x2 = matrix(rnorm(n * 10,mean=0,sd=.4),n,10)

simdata <- sim.survdata(N=n,T=365, censor=0.15, 
                        beta = c(0.4, 0.4, 0.5, 0.5, 0.5, 0, 0, 0, 0, 0,
                                 0.4, 0.4, 0.5, 0.5, 0.5, 0, 0, 0, 0, 0), 
                        X = cbind(x1,x2))

data <- simdata$data
attach(data)

setwd("~/StatisticalModelingandComputing/")
#save graphics output in pdf - saves graph(s) in working directory
 pdf(file="HW02_out.pdf")
```


```{r}
cormat <- round(cor(data[,1:20]),2)
cormat #the correlation coefficients among the predictors are small
corrplot(cormat,type="upper")
```


```{r}
model <- coxph(Surv(y, failed) ~ ., data=data)
model$coefficients ## model-estimated coefficients
summary(model)
```


```{r}
# Perform survival analysis using glmnet


# Perform survival analysis using glmnet
# assuming the predictors are in the first p columns and the response is in the last two columns
p <- ncol(data) - 2
p

xmatrix=as.matrix(data[,1:p])
show_fit=glmnet(xmatrix,Surv(y, failed),standardize=TRUE,lambda=seq(0,0.25,.001),alpha=1,family = "cox")
#print(show_fit)

plot(show_fit,label=TRUE)
plot(show_fit,xvar = "lambda",label=TRUE)
```


```{r}
#show the use of the CV metric "Partial Likelihood Deviance"
surv_model <- cv.glmnet(xmatrix,Surv(y, failed), family = "cox", type.measure = "deviance",alpha=1,nfolds = 10)
plot(surv_model)
print(surv_model)
```


```{r}
# Extract the optimal lambda value at lambda.1se
lambda_opt <- surv_model$lambda.1se

# Fit the model using the optimal lambda value
fit <- glmnet(xmatrix,Surv(y, failed), family = "cox", type.measure = "deviance",alpha=1,lambda = lambda_opt)
coef(fit)
```


```{r}
# Make predictions
predictions <- predict(fit, newx = xmatrix[,1:p], type = "response")

# Evaluate the model performance using concordance index (c-index)
c_index <- Cindex(predictions,Surv(y, failed))
c_index
```


```{r}
#show the use of CV metric "C" 
surv_model_C = cv.glmnet(xmatrix,Surv(y, failed), family = "cox", type.measure = "C", alpha=1,nfolds = 10)
plot(surv_model_C)
print(surv_model_C)
```


```{r}
# Extract the optimal lambda value at lambda.1se
lambda_opt <- surv_model_C$lambda.1se

# Fit the model using the optimal lambda value
fit_C <- glmnet(xmatrix,Surv(y, failed), family = "cox", type.measure = "C",alpha=1,lambda = lambda_opt)
coef(fit_C)
```


```{r}
# Make predictions
predictions <- predict(fit_C, newx = xmatrix[,1:p], type = "response")
# Evaluate the model performance using concordance index (c-index)
c_index <- Cindex(predictions,Surv(y, failed))
c_index
```


Assignment 2

a. Redo the simulation where you select 5 of the X coefficients among the binary X's to be 0 and 5 of the X coefficients among the normally distributed X's to be 0.

Note: just modify the line in the program:    beta = c(0.4, 0.4, 0.5, 0.5, 0.5, 0, 0, 0, 0, 0,
                                                                                         0.4, 0.4, 0.5, 0.5, 0.5, 0, 0, 0, 0, 0), 

Generate the text output and the graphical output to answer parts b - f below.


```{r}
#Simulate time-to-event data
set.seed(123)
n <- 1000

x1 = matrix(rbinom(n * 10, size = 1, prob = 0.3), n, 10)

x2 = matrix(rnorm(n * 10,mean=0,sd=.4),n,10)

simdata <- sim.survdata(N=n,T=365, censor=0.15, 
                        beta = c(0, 0.4, 0.5, 0, 0.5, 0, 0.4, 0.4, 0, 0.5, 
                                 0.5, 0, 0, 0, 0.4, 0, 0.5, 0.5, 0, 0),
                        X = cbind(x1,x2))

data <- simdata$data

```

```{r}
cormat <- round(cor(data[,1:20]),2)
cormat #the correlation coefficients among the predictors are small
corrplot(cormat,type="upper")
```
```{r}
model <- coxph(Surv(y, failed) ~ ., data=data)
model$coefficients ## model-estimated coefficients
summary(model)
```
```{r}
# Perform survival analysis using glmnet


# Perform survival analysis using glmnet
# assuming the predictors are in the first p columns and the response is in the last two columns
p <- ncol(data) - 2
xmatrix=as.matrix(data[,1:p])
show_fit=glmnet(xmatrix,Surv(data$y, data$failed),standardize=TRUE,lambda=seq(0,0.25,.001),alpha=1,family = "cox")
#print(show_fit)
```

```{r}
plot(show_fit,label=TRUE)
plot(show_fit,xvar = "lambda",label=TRUE) 
```
b. Did the Lasso proportional hazards model correctly select the 10 significant predictors using the partial likelihood deviance metric? Comment.

```{r}
#show the use of the CV metric "Partial Likelihood Deviance"
surv_model <- cv.glmnet(xmatrix,Surv(data$y, data$failed), family = "cox", type.measure = "deviance",alpha=1,nfolds = 10)
plot(surv_model)
print(surv_model)
```
```{r}
# Extract the optimal lambda value at lambda.1se
lambda_opt <- surv_model$lambda.1se

# Fit the model using the optimal lambda value
fit <- glmnet(xmatrix,Surv(data$y, data$failed), family = "cox", type.measure = "deviance",alpha=1,lambda = lambda_opt)
coef(fit)
```
All the 10 significant predictors were selected by using partial likelihood deviance without penalizing any coefficient to 0.

c. What is the C-index for the optimal lambda chosen to be Lambda.1se using the partial likelihood deviance metric?

```{r}
# Make predictions
predictions <- predict(fit, newx = xmatrix[,1:p], type = "response")

# Evaluate the model performance using concordance index (c-index)
c_index <- Cindex(predictions,Surv(data$y, data$failed))
c_index
```

The C-index using likelihood deviance metric is 0.6697602


d. Did the Lasso proportional hazards model correctly select the 10 significant predictors using Cross Validation with the C-index metric? Comment.

```{r}
#show the use of CV metric "C" 
surv_model_C = cv.glmnet(xmatrix,Surv(data$y, data$failed), family = "cox", type.measure = "C", alpha=1,nfolds = 10)
plot(surv_model_C)
print(surv_model_C)
```


```{r}
# Extract the optimal lambda value at lambda.1se
lambda_opt <- surv_model_C$lambda.1se

# Fit the model using the optimal lambda value
fit_C <- glmnet(xmatrix,Surv(data$y, data$failed), family = "cox", type.measure = "C",alpha=1,lambda = lambda_opt)
coef(fit_C)
```
Using CV all 10 significant predictors were selected without penalizing any coefficient to 0.

e.  What is the C-index for the optimal lambda chosen to be Lambda.1se using the C-index metric? 

```{r}
# Make predictions
predictions <- predict(fit_C, newx = xmatrix[,1:p], type = "response")
# Evaluate the model performance using concordance index (c-index)
c_index <- Cindex(predictions,Surv(data$y, data$failed))
c_index
```
The C-index using C-index metric is 0.6673913

f. Comment using just 1 sentence on the size of the correlations between pairs of the predictors.

Correlation between the pairs of predictors is small as its smaller than 1 and not very different than 0.6


```{r}
detach(data)
##------------------------------------------------------------------##
```

