---
output:
  pdf_document: default
  html_document: default
---


```{r}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("KMsurv")) install.packages("KMsurv")
if (!require("survival")) install.packages("survival")

library(KMsurv)
library(survival) 
```

```{r}
data(pneumon)

pneumon <- pneumon %>% arrange(desc(chldage))
head(pneumon)

```
#Creating censor variable in pneumon dataset
```{r}
censor=pneumon$chldage
censor=ifelse(censor >=12,0,1)
table(pneumon$chldage,censor)

head(pneumon)
pneumon['status'] <- ifelse(pneumon$chldage>=12,0,1)
head(pneumon)
```
# Part A
 a) Plot the Kaplan-Meier estimator for the survival (not having pneumonia) function for chldage (age at pneumonia). 

```{r}
fit.surv <- survfit(Surv(pneumon$chldage, pneumon$status) ~ 1)
plot(fit.surv, xlab = "Months",
    ylab = "Estimated Probability of not getting pneumonia",
    col = c(3,6))
```
#Give an estimate and a confidence interval for a newborn not having developed pneumonia at 6 months.

```{r}
summary(fit.surv, times = c(6))
```
From above summary at 6 months, the estimate of survival would be 0.801 with a confidence interval of 0.788(lower CI)-0.814(upperCI)

# Part b

Use the survfit function to stratify survival by poverty (Mother at poverty)

```{r}
fit.pov <- survfit(Surv(pneumon$chldage, pneumon$status) ~ pneumon$poverty)

fit.pov
```
  b) Generate the survival curves by poverty (1=yes 0=no) and perform the log-rank test
```{r}
plot (fit.pov , xlab = " Months ",
ylab = " Est. Prob. of Survival ", col = c(3,6))

logrank.test <- survdiff(Surv(pneumon$chldage, pneumon$status) ~ pneumon$poverty)
logrank.test

```

```{r}
#Poverty variable is an int in the dataset but can be used as a factor/categorical value, all calculations are done without converting into factors 

#pneumon$poverty <- as.integer(pneumon$poverty)

#fit.poverty1 <- survfit(Surv(pneumon$chldage, pneumon$status) ~ pneumon$poverty)
#summary(fit.poverty1)
#plot(fit.poverty1, xlab = "Months",
#     ylab = "Estimated Probability of not getting pneumonia using poverty as int", col = c(3,6))

#pneumon$poverty <- as.factor(pneumon$poverty)

#fit.poverty2 <- survfit(Surv(pneumon$chldage, pneumon$status) ~ pneumon$poverty)
#summary(fit.poverty2)
#plot(fit.poverty2, xlab = "Months",
#    ylab = "Estimated Probability of not getting pneumonia using poverty as factors", col = c(3,6))

```
#part c
#Fit the Cox proportional hazards model that includes only poverty as a covariate use ties='breslow' option

```{r}
#Fit the Cox proportional hazards model that includes only poverty as a covariate use ties='breslow' option

#without ties
fit.coxbasic <- coxph(Surv(pneumon$chldage, pneumon$status) ~ pneumon$poverty)
summary(fit.coxbasic) 
fit.coxbasic
```
#Is there a difference in survival between poverty levels (yes or no)


```{r}
#with ties

fit.coxties <- coxph(Surv(pneumon$chldage, pneumon$status) ~ pneumon$poverty, ties = "breslow")
summary(fit.coxties) 
fit.coxties

```
No, there is no difference hence no clear evidence for a difference in survival due to pneumonia between poverty groups with even using ties.

```{r}
#Fit a Cox proportional hazards model that includes 
# mthage, urban, alcohol, smoke, region, poverty, bweight, race, and education as predictors. use ties='breslow' option
fit.coxmultiplepred <- coxph(Surv(chldage, status) ~ mthage + urban + alcohol + smoke + region + poverty + bweight + race + education, data = pneumon, ties = "breslow")
summary(fit.coxmultiplepred) 
fit.coxmultiplepred


# List the significant predictors.

```
# Ans The significant predictors are mother age, smoke
